---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day

```{r}
dailysteps <- aggregate(steps ~ date, data=activity, FUN=sum)
barplot(dailysteps$steps, names.arg=dailysteps$date, xlab="date", ylab="steps")
```

2. Calculate and report the mean and median total number of steps taken per day

```{r}
mean(dailysteps$steps)
median(dailysteps$steps)
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
intervals <- aggregate(steps ~ interval, data=activity, FUN=mean)
plot(intervals, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
intervals$interval[which.max(intervals$steps)]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
